---
title: "A STATISTICAL RESERACH ON THE EFFECT OF MENTAL HEALTH ON STUDENTS CGPA DATASET"
author: "Mrs. Owoade Lola"
date: "7/23/2022"
output: github_document
---

## About the Data

This Data set was collected by a survey conducted by Google forms from University student in order to examine their current academic situation and mental health.

Therefore the indicators include:
-  Gender                                         (Categorical Variable)
-  Age                                            (Non-categorical Variable)
-  Course of study                                (Categorical Variable)
-  Year of study                                  (Categorical Variable)
-  CGPA                                           (Categorical Variable)
-  Marital status                                 (Categorical Variable)
-  Do you have Depression?                        (Categorical Variable)
-  Do you have Anxiety?                           (Categorical Variable)
-  Do you have Panic attack?                      (Categorical Variable)
-  Did you seek any specialist for a treatment?   (Categorical Variable)

### Dpendent Variable
- CGPA 

### Independent Variables
- Anxiety
- Depression
- Panic Attack

### Research Questions:
- Does Anxiety, Depression and Panic Attack has effect on the students CGPA?
- Does Age and Marital status affect CGPA

### Objectives:
- To describe and visualize the mental health of students with respect to their CGPA
- To determine the impact of Anxiety, Depression, and Panic attack on the students CGPA
- To investigate any relationship or association between gender, marital status and CGPA

### Hypothesis
 - H0: There is significance relationship between variables
 - H1: Not H0


#### Importing the dataset
```{r message=FALSE, warning=FALSE}
#reading the data
data <- read.csv(file.choose(), sep = ",", header = TRUE)
head(data)

attach(data)
```

from the data set, only Age is seen to be integer and quantitative, where all other variables are qualitative and categorical in nature.

## Data checking and Cleaning

```{r, message=FALSE, warning=FALSE}
summary(data)
```


# Exploratory Data Analysis (EDA)

## Univariate Analysis and  Graphical Representation

#### Importing required Packages
```{r, message=FALSE, warning=FALSE}
library(ggplot2)
library(dplyr)
```

### Gender Distribution
```{r echo=TRUE, fig.width=6, message=FALSE, warning=FALSE, paged.print=FALSE}
pi = data.frame(table(gender))
ggplot(pi, aes(x = "", y = Freq, fill = gender))+
  geom_bar(stat = "identity", width = 1, )+
  coord_polar("y", start = 0)+
  theme_bw()+
  labs(title = "Pie Chart Gender Distribution", x = " ", y = " ")
#A <- data$gender
#AB = data.frame(table(A))
#colours <- c("red","blue")
#percentlabels <- round(100*AB$Freq/sum(AB$Freq), 1)
#pielabel <- paste(c("No", "Yes"), " ", percentlabels, "%", sep="")
#pie(AB$Freq, main="Pie Chart for Gender", col=colours, labels = pielabel, cex=0.8)
# legend("topright", c("No", "Yes"),bty = "n",cex = 0.8, fill = colours)
```

### Descriptive Statistics of Age
```{r fig.height=7, message=FALSE, warning=FALSE}
library(dplyr)
su = data %>%
  summarize(n = n(),
            Min = min(Age, na.rm = T),
            Mean = mean(Age, na.rm = T),
            Median = median(Age, na.rm = T),
            Max = max(Age, na.rm = T),
            Standard_Dev = sd(Age, na.rm = T),
            Standard_Error = Standard_Dev/sqrt(n))
su
```
From the descriptive statistics of the students age in the table above, the average age of the students is 20, with the minimum and maximum age of 18 and 24. Also, the dispersion level of the studentd age is approximately 3, while the level at which the sample mean is likely to be, to the actual population mean is 0.25 respectively.

The Age is represented graphically with Box plot and Histogram Below 
### Boxplot of Age
```{r fig.height=7, message=FALSE, warning=FALSE}
boxplot(Age, col = "red", main = "Boxplot for Age", na.rm = T)
```

```{r}
hist(data$Age, col = "red", main = "Histogram of Age", xlab = "Age")
```


### Frequencies and Percentage of students CGPA
```{r}
d = data.frame(table(CGPA))
percentage = (d$Freq/sum(d$Freq) * 100)
Percentage = paste(percentage, "%")
CGP = data.frame(d,Percentage)
CGP
```

```{r echo=TRUE, fig.height=4, fig.width=6, message=FALSE, warning=FALSE, paged.print=FALSE}
library(ggplot2)
ggplot(data = data, aes(x=CGPA))+
  geom_bar(na.rm = TRUE, fill = "green")+
  theme_grey()+
  labs(title = "Bar plot for CGPA")
```


### Year of Study Distribution
```{r echo=TRUE, fig.height=4, fig.width=6, message=FALSE, warning=FALSE, paged.print=FALSE}
library(ggplot2)
ggplot(data = data, aes(x=year.of.Study))+
  geom_bar(na.rm = TRUE, fill = "red")+
  theme_grey()+
  labs(title = "Bar plot for Year of study")
```





### Marital status

```{r echo=TRUE, fig.height=4, fig.width=6, message=FALSE, warning=FALSE, paged.print=FALSE}
library(ggplot2)
pe = data.frame(table(Marital.status))
ggplot(pe, aes(x = "", y = Freq, fill = Marital.status))+
  geom_bar(stat = "identity", width = 1, )+
  coord_polar("y", start = 0)+
  theme_grey()+
  labs(title = "Pie chart for Marital Status", x = " ", y = " ")

```


### Frequencies of Anxiety, Depression, Panic attack and Treatment status 
```{r}
Responses = c("No", "Yes")
summ = data %>%
  summarize(Anxiety = table(Do.you.have.Anxiety.),
            Depression = table(Do.you.have.Depression.),
            Panic.Attack = table(Do.you.have.Panic.attack.),
            Treatment = table(Did.you.seek.any.specialist.for.a.treatment.))
describe = data.frame(Responses, summ)
describe
```

### Percentage of Anxiety, Depression, Panic attack and Treatment status
```{r}
most = describe %>%
  summarize( n = n(),
             Anxiety = Anxiety/sum(Anxiety)*100,
             Depression = Depression/sum(Depression)*100,
             Panic.Attack = Panic.Attack/sum(Panic.Attack)*100,
             Treatment = Treatment/sum(Treatment)*100)
most
```

### Do you have Depression?

```{r echo=TRUE, fig.height=4, fig.width=6, message=FALSE, warning=FALSE, paged.print=FALSE}
library(ggplot2)
pw = data.frame(table(Do.you.have.Depression.))
ggplot(pw, aes(x = "", y = Freq, fill = Do.you.have.Depression.))+
  geom_bar(stat = "identity", width = 1, )+
  coord_polar("y", start = 0)+
  theme_grey()+
  labs(title = "Pie chart for Depression Status", x = " ", y = " ")

```

### Do you have Anxiety?

```{r echo=TRUE, fig.height=4, fig.width=6, message=FALSE, warning=FALSE, paged.print=FALSE}
library(ggplot2)
pw = data.frame(table(Do.you.have.Anxiety.))
ggplot(pw, aes(x = "", y = Freq, fill = Do.you.have.Anxiety.))+
  geom_bar(stat = "identity", width = 1, )+
  coord_polar("y", start = 0)+
  theme_grey()+
  labs(title = "Pie chart for Anxiety Status", x = " ", y = " ")

```


### Do you have Panic Attack?

```{r echo=TRUE, fig.height=4, fig.width=6, message=FALSE, warning=FALSE, paged.print=FALSE}
library(ggplot2)
pw = data.frame(table(Do.you.have.Panic.attack.))
ggplot(pw, aes(x = "", y = Freq, fill = Do.you.have.Panic.attack.))+
  geom_bar(stat = "identity", width = 1, )+
  coord_polar("y", start = 0)+
  theme_grey()+
  labs(title = "Pie chart for Panic attack Status", x = " ", y = " ")

```


### Did you seek any specialist for a treatment?
```{r echo=TRUE, fig.height=4, fig.width=6, message=FALSE, warning=FALSE, paged.print=FALSE}
library(ggplot2)
pw = data.frame(table(Did.you.seek.any.specialist.for.a.treatment.))
ggplot(pw, aes(x = "", y = Freq, fill = Did.you.seek.any.specialist.for.a.treatment.))+
  geom_bar(stat = "identity", width = 1, )+
  coord_polar("y", start = 0)+
  theme_grey()+
  labs(title = "Pie chart for specialist treatment status", x = " ", y = " ")

```




## MULTIVARIATE DATA ANALYSIS AND VISUALIZATIONS

### Comparative analysis of Average Age with respect to gender
```{r}
plotdata  <- data %>%
  group_by(gender) %>%
  summarize(n = n(),
            mean = mean(Age, na.rm=T),
            sd = sd(Age, na.rm = T),
            se = sd/sqrt(n))
plotdata

```

### Comparative column plot of mean Age with respect to Gender
```{r}
plotme  <- data %>%
  select(gender, Age)%>%
  group_by(gender) %>%
  summarize(age = mean(Age, na.rm = T)) %>%
  ggplot(aes(x = gender, y = age))+
  geom_col(fill = "brown")+
  labs(title = "Column plot of mean age, grouped by gender")

plotme
```

### Comparative Bar plot of Gender with respect to CGPA
```{r}
  ggplot(data, aes(x = gender, fill = CGPA))+
  geom_bar(position = "dodge")
```

### Comparative Bar plot of CGPA with respect to Gender
```{r}
  ggplot(data, aes(x = CGPA, fill = gender))+
  geom_bar(position = "dodge")
```

### Comparative Bar plot of study year with respect to CGPA
```{r}
  ggplot(data, aes(x = CGPA, fill = year.of.Study))+
  geom_bar(position = "dodge")
```


### Comparative Bar plot of Marital status with respect to CGPA 
```{r}
  ggplot(data, aes(x = CGPA, fill = Marital.status))+
  geom_bar(position = "dodge")
```

### Comparative Bar plot of Anxiety status with respect to CGPA
```{r}
  ggplot(data, aes(x = CGPA, fill = Do.you.have.Anxiety.))+
  geom_bar(position = "dodge")
```


### Comparative Bar plot of Depression with respect to CGPA
```{r}
  ggplot(data, aes(x = CGPA, fill = Do.you.have.Depression.))+
  geom_bar(position = "dodge")
```


### Comparative Bar plot of Panic attack with respect to CGPA
```{r}
  ggplot(data, aes(x = CGPA, fill = Do.you.have.Panic.attack.))+
  geom_bar(position = "dodge")
```

### Comparative Bar plot of Treatment status with respect to CGPA
```{r fig.width=10}
  ggplot(data, aes(x = CGPA, fill = Did.you.seek.any.specialist.for.a.treatment.))+
  geom_bar(position = "dodge")
```

### Comparative Bar plot of Age with respect to CGPA
```{r}
  ggplot(data, aes(x = Age, fill = CGPA))+
  geom_bar(position = "dodge", na.rm = T)
  #geom_text(aes(label=round( ,1)), hjust = -.1)
```





# INFERENCIAL STATISTICS

## Investigating  the impact of Anxiety on  students CGPA

```{r}
a = table(CGPA, Do.you.have.Anxiety.)
a
```
```{r message=FALSE, warning=FALSE}
chisq.test(a)
```

```{r}
b = table(CGPA, Do.you.have.Depression.)
b
```
```{r message=FALSE, warning=FALSE}
chisq.test(b)
```

```{r}
c = table(CGPA, Do.you.have.Panic.attack.)
c
```
```{r message=FALSE, warning=FALSE}
chisq.test(c)
```

## To investigate any relationship or association between gender, marital status and CGPA

```{r}
d = table(CGPA, gender)
d
```

```{r message=FALSE, warning=FALSE}
chisq.test(d)
```


```{r}
e = table(CGPA, Marital.status)
e
```

```{r message=FALSE, warning=FALSE}
chisq.test(e)
```


# FINDINGS AND CONCLUSIONS

- From the result, female seems to be higher than male with frequency ratio of 75:26, and the students CGPA seems to be depending on age as female students are seen to have the highest CGPA in each CGPA grades, with same averages of ages (20years). 
- Also, CGPA seems to be having some kind of relationship with students year of study, as students in first years shows to have the highest CGPA, while those in the forth level has the lowest counts of grade point, and the CGPA counts decreases over the study years especially in the Upper level and Distinction grade points.
- Marital staus also shows a positive impact on CGPA as single students dominated Upper credit and Distinction grape point, though few married students also dominated these points, but not even close to that of single students.
- Amxiety as well have an impact on CGPA as students with no anxiety are far more than those with anxiety in the grade points.
- Depression can also determine students CGPA as students with no depression cases also counted higher than those with depressions in the grade points, except in lower credit, and same thing goes for Panic attack as well.
- Also, Age proves to determine CGPA with the highest distinction and upper credit counts falling under the lowest age of 18.


Note: However, we cannot make much conclusions  with these visualization results as the sample size is is small, and the gender distribution counts is too far from each other.


- From the chisquare test, we can conclude that anxiety has no significance relationship with CGPA, neither does Depression nor Panic attack has any significance relationship with CGPA.
- Gender and Marital status also tested to have no statistical significance with CGPA


- Therefore , we can conclude that none of depression, anxiety, panic attack, gender or marital status cannot determine CGPA, and students can pass and have good grades despite experiencing anxiety, depression, panic attack, being any of the gender or even being single or married.

